Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-23 paper/paper.tex
done my pass on section 3. moving to section 4.

Second, \tname{} uses the benign traffic to mitigate the overfitting
issue. In this step, \tname{} searches for alternative rules that
preserve the invariant to capture the positive malicious traffic but
not capture the negative benign traffic. It uses the rule obtained in
the previous step as seed to bootstrap the search for alternative
plausible rules, \ie{}, rules that satisfy the aforementioned
invariant. The search is designed to generate rules containing subsets
of the options from the seed rule. Considering this example, \tname{}
produces a total of \Fix{XXXXX} plausible rules at the end of the
search.\Mar{Technically, there are $32 (=2^5)$ subsets for a set with 5
  elements. So, it is important to indicate why Table 3 shows sooo
  many plausible rules and that this is is a very simple rule (for illustration).}

Finally, \tname{} ranks plausible rules according to heuristics
learned from a database of existing rules. The intuition is that new
rules should share characteristics with pre-existing rules. For
example, one heuristic function used by \tname{} leverages the amount
of options a rule typically has. The score of a rule, given by this
function, is proportional to the popularity of the number of options
declared in rules of a given
protocol. Figure~\ref{fig:distribution-contents} shows the
distribution of the number of content options for the database of
rules we used\Mar{reference} \Luc{essa distribuicao n eh utilizada 
para pontuar as regras desse exemplo, n seria mais interessante citar outra?
poderia ser a de tamanhos das regras.}. Note that rules with 1 to 3 options 
are more common \Luc{1 to 3 \textbf{content} options}. At the end of the 
process, \tname{} ranks the golden rule at the top of the ranking 
including \Fix{XXXXXX} other plausible rules.

\Mar{add short par indicating that other problems can be much more
  complex than this...}

@@

\tname\ works as follows. First, it produces a potentially
overspecified rule that captures the malicious traffic, provided on
input in the form of pcap files~\cite{pcap}. In this step,
\tname\ detects rule options directly from the input messages (see
Section~\ref{sec:rules-and-packets}). As only malicious traffic is
considered in this step, the rule produced is likely overspecified,
\ie{}, it includes more options than the golden rule (see
Section~\ref{sec:suri-metas-coverage}). Consequently, the use of that
rule could lead to false negatives as not all manifestations of the
same attack are identical. Second, \tname{} generates plausible rules
that are similar to the seed rule produced in the previous step. It is
worth noting that any subset of the original set of options captures
the malicious traffic. \tname{} leverages that observation to search
for rules containing a proper subset of rules defined in the original
seed option. As the generation procedure discards options from the
initial rule, the rules \tname{} creates satisfies the property of
capturing malicious traffic. However, the goal of \tname\ is to
produce rules that capture \emph{only} malicious traffic. To achieve
that goal, \tname{} uses benign traffic to reduce the likelihood the
\nids\ will report false positives on the synthesized rule. The
minimization step uses a population-based gradient descent-like search
to find all plausible rules derived from the seed rule, obtained in
the previous step. Third, \tname{} uses several heuristic functions
derived from a dataset of rules to rank the rules generated in the
previous step.

@@

\tname{} leverages public datasets~\cite{tcpreplay} of benign traffic
to guide the search for more general rules, hence addressing the
overffiting issue. To that end, \tname{} \emph{minimizes} the rules
produced in the previous step, preserving the invariant of \emph{only}
capturing the positive malicious traffic. We say that a rule is
minimal if discarding any option results in capturing negative (\ie{},
benign) traffic.  The minimization procedure we propose in this step
builds on the observation that all options from the rule produced in
the previous step are satisifed by the positive traffic. Consequently,
any subset of these options also captures the positive traffic. The
procedure iteratively discards options from the candidate rule until
it detects that the rule is minimal, \ie{}, it finds that no other
option can be removed. The procedure uses negative traffic to guide
the search.

@@

\vspace{2ex}
\noindent\textbf{Limitations.} It is worth noting that although
\tname{} is accurate for generating rules that discrimate positive and
negative traffic, the quality of these rules is limited by the quality
of the training data it uses. For the positive traffic, \tname{} only
requires one sample, but can benefit of several samples as the example
from Section~\ref{sec:dos} illustrated. The same applies for the
negative traffic. In this case, however, \tname{} can use public
datasets available online.

@@

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity between a
synthesized and a human-created rule is the size of the synthetic
rule, \ie{}, the number of options it contains. Informally, rules with
popular sizes score high whereas inpopular rules score
low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes for our rule dataset. The x-axis shows the number of
options included in a rule (up to 10) while the y-axis shows the
number of rules with the corresponding number of options. Note that
there are 11 rules with no options. These rules only use the header to
signal a traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

@@

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\noindent
, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our rule dataset. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2).

@@

%% we observed---by analyzing
%% the dataset of existing rules (see
%% Section~\ref{sec:dataset-benign})---that these rules often indicate
%% what part of the message the content string should be checked
%% against. For example,
%% For example, the
%% creator of the rule may want to indicate that a specific string be
%% present in the header of an HTTP request message. Content
%% \emph{modifiers} serve this purpose.
%% Suricata supports content modifiers for tcp and http. \Luc{Eh possivel utilizar os modificadores em qualquer protocolo. Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo. O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós. Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso. Se necessário posso explicar.} Given that
%% content options often use modifiers\Mar{assumindo que eh
%%   verdade. confirmar estat. solicitada acima...},

\tname{} leverages the fact that these modifiers are commonly used to
define rules as to better rank the rules it synthesizes. More
precisely, it builds on the observation that strings acessible in
positions associated with content modifiers (\eg{}, GET, Host,
Connection) are more likely to produce useful rules. \tname\ uses the
frequency of content modifiers to discriminate rules that are more or
less likely to reflect actual rules. For example, historical data
indicates that content options that use the \CodeIn{http\_connection}
modifier are much less likely than content options that use the
modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our dataset.

@@

%% \Mar{Qual o rationale para considerar as varias
%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes? Não eh o caso. A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada. Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content. Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}

@@

\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
traffic for all protocols covered by Suricata (see
Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
benign traffic. The Bigflows.pcap dataset is part of the
TcpReplay~\cite{tcpreplay} open source project. This dataset includes
real network traffic of a busy private network access point to the
Internet. According to the TcpReplay web site ``This capture is much
larger and has a smaller average packet size than the previous capture
(smallFlows.pcap). It also has many more flows and different
applications.''. In addition to the Bigflows.pcap dataset, we also
used datasets of normal traffic made available by Stratosphere
Lab~\cite{stratosphere-normal}.

@@

%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood. Os datasets foram encontrados na internet.}

>>>>

Second, \tname{} uses the benign traffic to mitigate the overfitting
issue. In this step, \tname{} searches for alternative rules that
preserve the invariant to capture the positive malicious traffic but
not capture the negative benign traffic. It uses the rule obtained in
the previous step as seed to bootstrap the search for alternative
plausible rules, \ie{}, rules that satisfy the aforementioned
invariant. The search is designed to generate rules containing subsets
of the options from the seed rule. Considering this example, \tname{}
produces a total of \pingscanplausible{} plausible rules at the end of
the search.  In the worst case, \tname{} produces $2^n$ plausible
rules, where $n$ is the number of options contained in the inferred
rule ($n=5$, in this case). Rules are discared in the search process
as they are found to be implausible.

%% \Mar{Technically, there are $32 (=2^5)$ subsets for a set
%%   with 5 elements. So, it is important to indicate why Table 3 shows
%%   sooo many plausible rules and that this is is a very simple rule
%%   (for illustration).}

Finally, \tname{} ranks plausible rules according to heuristics
learned from a database of existing rules. The intuition is that new
rules should share characteristics with pre-existing rules. For
example, one heuristic function used by \tname{} leverages the
\emph{size of a rule} for that. The score of a rule given by such a
function is proportional to the popularity of the number of options
declared in rules of a given
protocol. Figure~\ref{fig:distribution-rule-size} shows the
distribution of the number of content options associated with rules
from a public data set~\cite{emerging-threats-open}. Note that rules
with 2 to 6 options are more common. At the end of this step, \tname{}
ranks the golden rule at the top of the ranking, including
\pingscanplausible{} rules. It is worth noting that
\tname\ consistently ranks the golden rule very high in most of the
cases even when the seed rule, produced at the first step, contains a
large number of options (see Table~\ref{table:results}).

@@

\tname\ works as follows. First, it produces a potentially
overspecified rule that captures the malicious traffic, provided on
input in the form of pcap files~\cite{pcap}. In this step,
\tname\ detects rule options directly from the input messages (see
Section~\ref{sec:rules-and-packets}). As only malicious traffic is
considered in this step, the rule produced is likely overspecified,
\ie{}, it includes more options than the golden rule (see
Section~\ref{sec:suri-metas-coverage}). Consequently, the use of that
rule could lead to false negatives as not all manifestations of the
same attack are identical. Second, \tname{} generates plausible rules
that are similar to the seed rule produced in the previous step. It is
worth noting that any subset of the original set of options captures
the malicious traffic. \tname{} leverages that observation to search
for rules containing a proper subset of rules defined in the original
seed option. As the generation procedure discards options from the
initial rule, the rules \tname{} creates satisfies the property of
capturing malicious traffic. However, the goal of \tname\ is to
produce rules that capture \emph{only} malicious traffic. To achieve
that goal, \tname{} uses benign traffic to reduce the likelihood the
\nids\ will report false positives on the synthesized rule. The
minimization step uses a population-based gradient descent-like search
to find all plausible rules derived from the seed rule, obtained in
the previous step. Third, \tname{} uses several heuristic functions
derived from a data set of rules to rank the rules generated in the
previous step.

@@

\tname{} leverages public data sets~\cite{tcpreplay} of benign traffic
to guide the search for more general rules, hence addressing the
overffiting issue. To that end, \tname{} \emph{minimizes} the rules
produced in the previous step, preserving the invariant of \emph{only}
capturing the positive malicious traffic. We say that a rule is
minimal if discarding any option results in capturing negative (\ie{},
benign) traffic.  The minimization procedure we propose in this step
builds on the observation that all options from the rule produced in
the previous step are satisifed by the positive traffic. Consequently,
any subset of these options also captures the positive traffic. The
procedure iteratively discards options from the candidate rule until
it detects that the rule is minimal, \ie{}, it finds that no other
option can be removed. The procedure uses negative traffic to guide
the search.

@@

\vspace{2ex}
\noindent\textbf{Limitations.} It is worth noting that although
\tname{} is accurate for generating rules that discrimate positive and
negative traffic, the quality of these rules is limited by the quality
of the training data it uses. For the positive traffic, \tname{} only
requires one sample, but can benefit of several samples as the example
from Section~\ref{sec:dos} illustrated. The same applies for the
negative traffic. In this case, however, \tname{} can use public
data sets available online.

@@

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity between a
synthesized and a human-created rule is the size of the synthetic
rule, \ie{}, the number of options it contains. Informally, rules with
popular sizes score high whereas inpopular rules score
low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes for our rule data set. The x-axis shows the number of
options included in a rule (up to 10) while the y-axis shows the
number of rules with the corresponding number of options. Note that
there are 11 rules with no options. These rules only use the header to
signal a traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

@@

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\noindent
, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our rule data set. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2).

@@

%% we observed---by analyzing
%% the data set of existing rules (see
%% Section~\ref{sec:data set-benign})---that these rules often indicate
%% what part of the message the content string should be checked
%% against. For example,
%% For example, the
%% creator of the rule may want to indicate that a specific string be
%% present in the header of an HTTP request message. Content
%% \emph{modifiers} serve this purpose.
%% Suricata supports content modifiers for tcp and http. \Luc{Eh possivel utilizar os modificadores em qualquer protocolo. Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo. O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós. Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso. Se necessário posso explicar.} Given that
%% content options often use modifiers\Mar{assumindo que eh
%%   verdade. confirmar estat. solicitada acima...},

\tname{} leverages the fact that these modifiers are commonly used to
define rules as to better rank the rules it synthesizes. More
precisely, it builds on the observation that strings acessible in
positions associated with content modifiers (\eg{}, GET, Host,
Connection) are more likely to produce useful rules. \tname\ uses the
frequency of content modifiers to discriminate rules that are more or
less likely to reflect actual rules. For example, historical data
indicates that content options that use the \CodeIn{http\_connection}
modifier are much less likely than content options that use the
modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our data set.

@@

%% \Mar{Qual o rationale para considerar as varias
%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes? Não eh o caso. A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada. Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no data set.}
%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content. Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}

@@

\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
traffic for all protocols covered by Suricata (see
Figure~\ref{fig:distribution-rules-protocol}). We used two data sets of
benign traffic. The Bigflows.pcap data set is part of the
TcpReplay~\cite{tcpreplay} open source project. This data set includes
real network traffic of a busy private network access point to the
Internet. According to the TcpReplay web site ``This capture is much
larger and has a smaller average packet size than the previous capture
(smallFlows.pcap). It also has many more flows and different
applications.''. In addition to the Bigflows.pcap data set, we also
used data sets of normal traffic made available by Stratosphere
Lab~\cite{stratosphere-normal}.

@@

%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser
%% compostos por milhares de pacotes, um dos data sets q encontrei
%% possui mais de 500k pacotes de flood. Os data set  s foram encontrados na internet.}
